---
output: pdf_document
---
# Part 2

```{r, echo=F, message=F}
set.seed(1)
library(igraph)
```

Specify a stochastic block matrix:
We believe fraudsters don't often trade with themselves or with honest, but mostly with accomplices.  
Accomplices try to look normal by trading a lot with honest people, and occasionally with fraudsters. 
Honest trade with accomplices and themselves, but rarely with fraudsters.  
```{r}
# stochastic block matrix 
# our graph will have three communities
eps <- .021 # epsilon

M = matrix(c( 
  eps,     1-2*eps,       eps,
  1-2*eps,   2*eps,  .5-2*eps,
  eps,    .5-2*eps,  .5-2*eps
), 
nrow = 3)
M
```
These are symmetric:

```{r}
t(M) == M
```

Let's build our graph
```{r}
# sample a random graph
# 100 nodes grouped into 3 communities
num.fraudsters <- num.accomplices <- 5
num.honest <- 90
rg = 
  sample_sbm(num.fraudsters + num.accomplices + num.honest, # number of nodes  
             pref.matrix = M,      
             block.sizes = c(num.fraudsters,
                             num.accomplices,
                             num.honest),  # how many nodes belong to each community 
             loops = F,            # no loops (vertex that connects to itself)
             directed = F          # we want an undirected graph 
  ) 

# membership vector used to color vertices 
membership_vector = c(rep(1, num.fraudsters), 
                      rep(2, num.accomplices), 
                      rep(3, num.honest))

plot_layout = layout.fruchterman.reingold(rg)
plot(rg, 
     vertex.color=membership_vector,
     layout = plot_layout)
```
Our plot shows that the fraudsters are separate from the rest, but it's hard to identify the honest from accomplices, which is the point of accomplices.  

```{r}
library(lda)

result =
  mmsb.collapsed.gibbs.sampler(get.adjacency(rg),  # the adjacency matrix
                               K = 3,              # 3 groups
                               num.iterations=10000,    
                               alpha = 0.1,
                               burnin = 500L,
                               beta.prior = list(1, 1))

# this matrix tells us for each vertex what is the probability that it belongs to 
# one of the K communities
memberships = with(result, t(document_sums) / colSums(document_sums))
head(memberships,20)
```
Many of these are clearly predicted to be from one community versus the others, indicating that we think they likely belong to a given community.  

We expect the first five to be fraudsters, group 1:
```{r}
head(memberships[1:5,])
```
We see that they are classified together.  

And the expected accomplices:
```{r}
memberships[6:10,]
```
These are mainly classified together as well. 

The honest ones?
```{r}
memberships[11:30,]
mean(memberships[11:100,1])
mean(memberships[11:100,2])
mean(memberships[11:100,3])
```
The honest ones look to be grouped as well. 

```{r}
# the estimate of the stochastic block matrix M 
ratio = with(result, blocks.pos / (blocks.pos + blocks.neg))
ratio

# actual M
M
```
Community assignments got shifted around, but we classified consistently.  If we were to rearrange one of these matrices so the groupings were consistent, we might see them line up well.  